---
title: "13 - Clustering"
#output: rmarkdown::html_vignette
output:
  pdf_document:
    fig_crop: yes
vignette: >
  %\VignetteIndexEntry{case_study}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  #fig.height = 2,
  #fig.asp=1,
  fig.align = "center",
  out.extra='keepaspectratio',
  out.width ="\\textwidth"
)
```

```{r setup}
library(clustering)
```

```{r plot_helper, include=FALSE, cache=TRUE}
plot2d <- function (data, title="Untitled", xlabel=NULL, ylabel=NULL ) {
  plot(data[1,], data[2,], cex.axis=.75, pch=1, cex=.5, col=rgb(0,0,0,.5), main=title)
  if (!missing(xlabel)) mtext(side=1, line=2, xlabel);
  if (!missing(ylabel)) mtext(side=2, line=2, ylabel);
}
```

# Spectral Clustering

intro text here

## Perfect Projections



```{r, cache=TRUE, echo=FALSE, fig.asp=.6}
par(mfrow=c(1,2), mar=c(3,3,2,1))

set.seed(1)

X <- clustering::generate_nested_2d_training_data(400)
plot2d(X, title="Input Data",xlabel="", ylabel="")

X_projections <- spectral_clustering(X, 2, gaussian_kernel_with_fixed_gamma(30));
plot2d(X_projections, title="Projections for gamma = 30",xlabel="", ylabel="")
```

## Retrieve Clusters





# K-Medoids Algorithm

The k-medoids algorithm is a clustering approach related to k-means clustering. It is used to partition a data set into k clusters. Each cluster is represented by one medoid. A medoid can be defined as the point in the cluster, whose dissimilarities with all the other points in the cluster is minimum. The k-medoids algorithm requires the user to specify the number of clusters k

## Example

```{r, cache=TRUE, echo=FALSE, fig.asp=.6}

X <- matrix(c(8,7,3,7,4,9,9,6,8,5,5,8,7,3,8,4,7,5,4,5), nrow = 2)
plot2d(X, title="Input Data",xlabel="", ylabel="")

#clustering::plot_clustered_2d_data(k_medoids(X, 2), show_noise = FALSE)
```
