---
title: "R Course Project"
subtitle: "Clustering"
output: ioslides_presentation
vignette: >
  %\VignetteIndexEntry{presentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(clustering)
```

# Introduction


## The Goal of Clustering

- given a set of input data
- try to identify related vectors
- groups of related vectors are called a ***cluster***

```{r, echo=FALSE, fig.asp=.55, warning=FALSE}
par(mfrow=c(1,2), mar=c(3,3,2,1))

set.seed(188)
n <- 50
x <- cbind(
  x = runif(5, 0, 2) + rnorm(n, sd = 0.05),
  y = runif(5, 0, 2) + rnorm(n, sd = 0.05)
)
x <- t(x)
d <- c(x, .6, .75, .8, 1.3, 0, 1.6, .6, 1, .9, .9, .5, 1.5, 1, .4, .1, .6);
x <- matrix(d, nrow=2);

y <- x
attr(y, "cluster") <- rep(-1, n)
clustering::plot_clustered_2d_data(y,show_noise = TRUE);
clustering::plot_clustered_2d_data(DBSCAN(x, .1, 7));
```

## Input Data

- data is saved in matrices
- columns are vectors
- $\text{data} \subset \mathbb{R}^{\text{nrow(data)}}$
- $\#\text{data} = ncol(data)$

```{r, echo=FALSE, fig.cap="Sample data: Five two-dimensional Vectors"}
print(generate_2d_cluster(5))
```

## Output Data

- algorithms return input data
- cluster information saved in the attribute `"cluster"`
- non-positive cluster values denote noise, e.g. `-1`

```{r, echo=FALSE}
X <- generate_2d_cluster(5)
attr(X, "cluster") <- c(1,-1,2,2,1)
print(X)
```


# Clustering Example

## Spectral Clustering

- algorithm does generated projections of the vectors
- does not directly produce a clustering
- by running another algorithm on the projections clusters can be found

<!-- Note: - colors are not visible to the algorithm -> only for visualization purposes -->
<!--       - k-medoids algorithm used to find clusters -->


```{r, fig.asp=.35, cache=TRUE, echo=FALSE}
par(mfrow=c(1,3), mar=c(3,3,2,1))

set.seed(1)

X <- clustering::generate_nested_2d_training_data(400)
plot(X[1,], X[2,], xlab="", ylab="", col=c("darkred", "darkblue")[attr(X, "innerOrOuter")])
title("Nested Clusters")

data <- clustering::spectral_clustering(X, k=2)
plot(data[1,], data[2,], col=c("darkred", "darkblue")[attr(X, "innerOrOuter")], xlab="", ylab="", main="Generated Projections")

cData <- clustering::spectral_clustering(X, k=2, mercer_kernel=gaussian_kernel_with_fixed_gamma(30), clustering=k_medoids, 2);
plot_clustered_2d_data(cData)
title("Resulting Clustering")
```

## TODO: another clustering example


# The Package
##
3. Paket vorstellen,
Was erwarten die Clustering Funktionen f체r Argumente
Was geben die f체r Ausgaben
Liste aller Funktionen
Darauf eingehen, das die alle ungef채hr dieselben Argumente brauchen


# Live Demonstration
##
4. Live-Beispiele
nur mit der Standardfunktionalit채t
(zu 4.: Fehlerausgaben funktionieren usw)
(zu 4.: Dokumentation)


# Special Functionality
##
5. Besonderen Funktionen
(Hierarchical Clustering: Dendrogramm, Voronoi-Zellen bei K-means, Optics die extra-Graphen)

# Real Data Example
##
6. Beispiel mit richtigen Daten
