---
title: "13 - Clustering: Functionality and Usage"
output: pdf_document
#output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{13 - Clustering}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(clustering)
```

# Clustering Alforithms

The package contains implementations of the following algorithms:

* k-Means
* k-Mediods
* Hierarchical Clustering
* Spectral Clustering
* DBSCAN
* OPTICS

In this vignette we demonstrate how these functions are used, the resulting data, and additional functionality of the package. A case study demonstrating the clustering methods themselves can be found in a separate document.

# General Structure of Vector Data

Every function in this package expects/uses matrices whose columns are the input/output vectors, e.g. a collection of two-dimensional vectors is stored in a matrix `m` with `nrow(m)=2`.

```{r, echo=FALSE}
print(generate_2d_cluster(5))
```
Clusters produced by the clustering algorithms will be stored in the `"cluster"` attribute of the output data. A negative value always denotes noise.
```{r, echo=FALSE}
X <- generate_2d_cluster(5)
attr(X, "cluster") <- c(1,-1,2,2,1)
print(X)
```


# Helper Functions for Demonstration and Testing Purposes

The package exports a few functions to generate and plot two-dimensional clusters which may be used to familiarize oneself with the functionality of and the difference between the clustering methods.

```{r, eval=FALSE}
generate_2d_cluster(n, center=c(0, 0))
```
Generates a single cluster of `n` vectors in a radius of 1 around a given `center` point.

```{r, eval=FALSE}
generate_nested_2d_training_data(n)
```
Generates two nested clusters of `n` vectors. In addition, the returned data contains an extra attribute specifying to which cluster a given vector belongs. This is not used by any clustering algorithm but may be used to color data in a plot or to check the quality of a clustering algorithms result.

The following figure demonstrates how to use the latter function:

```{r, fig.width=4, fig.height=4}
data <- generate_nested_2d_training_data(400)
plot(data[1,], data[2,], col=c("darkred", "darkblue")[attr(data, "innerOrOuter")], xlab="", ylab="")
```

```{r, eval=FALSE}
plot_clustered_2d_data(data, point_size=.75, show_noise=TRUE, show_legend=FALSE, hide_axis_text=FALSE)
```
This function may be used to conveniently plot data clustered by this package's clustering algorithms. Only the first two dimensions are plotted, extra dimensions are ignored.

# K-Medoids Algorithm

The k-medoids algorithm is a clustering approach related to k-means clustering. It is used to partition a data set into k clusters using medoids. The k-medoids algorithm requires the user to specify the number of clusters k.

```{r}
data <- matrix(c(8,7,3,7,4,9,9,6,8,5,5,8,7,3,8,4,7,5,4,5), nrow = 2)
k_medoids(data, 2)
```

# Spectral Clustering

```{r, eval=FALSE}
spectral_clustering(data, k=1, mercer_kernel=gaussian_kernel_with_fixed_gamma(15), clustering=NULL, ...)
```
The spectral clustering algorithm produces projections of the input data into lower dimensions, by passing a `k` value the target dimension may be specified. Should the given `k` be larger than the input's dimension the function will report an error. The Mercer-Kernel used during the generation of the output projections may also be customized. Since the spectral clustering algorithm itself only projects the input data, a second algorithm has to be employed. This can either be done manually, using the data returned by the `spectral_clustering` function or by passing a clustering algorithm as the `clustering` argument. Any further arguments will be passed into the given clustering algorithm. If the `clustering` argument is set, the function will return clustered data.

```{r, cache=TRUE, echo=FALSE}
par(mfrow=c(1,2), mar=c(3,3,2,1))

set.seed(1)

X <- clustering::generate_nested_2d_training_data(400)
data <- clustering::spectral_clustering(X, k=2)
plot(data[1,], data[2,], col=c("darkred", "darkblue")[attr(X, "innerOrOuter")], xlab="", ylab="", main="Projections for gamma = 30")

cData <- clustering::spectral_clustering(X, k=2, mercer_kernel=gaussian_kernel_with_fixed_gamma(30), clustering=kmean, 2);
attr(X, "cluster") <- attr(cData, "clusters")
attr(cData, "cluster") <- attr(cData, "clusters")
plot_clustered_2d_data(cData)
```


